```{r}
a <- c(1, 3, 2)
t(a)
3*a
```

```{r}
b <- c(2, 8, 9)
a+b
```

```{r}
sum(a * b) #the dot product
```

```{r}
A <- matrix(c(1, 3, 2, 2, 8, 9), ncol = 3)
A
7 * A
t(A)
sum(diag(A)) #trace of a matrix
```

```{r}
A2 <- matrix(c(1, 3, 2, 2, 8, 9), ncol = 3, byrow = T)

B <- matrix(c(5, 8, 3, 4, 2, 7), ncol = 3, byrow = T)
matrix(0, nrow = 2, ncol = 3)
matrix(1, nrow = 2, ncol = 3)
diag(c(1, 2, 3))
```

```{r}
 A <- matrix(c(1, 3, 2, 2, 8, 9), ncol = 2)
 B <- matrix(c(5, 8, 4, 2), ncol = 2)
 A %*% B # Matrix Product
```
 
```{r}
 #inverse of matrix
A <- matrix(c(1, 3, 2, 4), ncol = 2, byrow = T)
B=solve(A) # Gets inverse of A
```

```{r}
#solving a linear system
  A <- matrix(c(1, 2, 3, 4), ncol = 2)
  b <- c(7, 10)
  x <- solve(A) %*% b
  x
```

```{r}
library(MASS) 
source("https://foxweb.marist.edu/users/duy.nguyen2/hdr2d.r")

#### Simulate multivariate normal vector
# Gives a random vector
# r -> random; mv -> multivariate
# norm -> normal
rmvnorm<-
function(n,mu,Sigma) {
  p<-length(mu)
  res<-matrix(0,nrow=n,ncol=p)
  if( n>0 & p>0 ) {
    E<-matrix(rnorm(n*p),n,p)
    res<-t(t(E%*%chol(Sigma)) + c(mu))
  }
  res
 }
```


```{r}
#### Log density of the multivariate normal distribution
ldmvnorm<-function(y,mu,Sig){  # log mvn density
  c(  -(length(mu)/2)*log(2*pi) -.5*log(det(Sig)) -.5*
        t(y-mu)%*%solve(Sig)%*%(y-mu)   )  
}
```

```{r}
#### Figure 7.1
#pdf("fig7_1.pdf",family="Times",height=1.75,width=5)

mu<-c(50,50)
sig<-c(8,12)
ng<-50
par(mfrow=c(1,3),mar=c(2.75,2.75,.5,.5),mgp=c(1.70,.70,0))
y<-seq(20,80,length=ng)
set.seed(1)
```

```{r}
for( rho in c(-.5,0,.5)) {

  Sigma<- matrix( c(1,rho,rho,1),2,2) * outer(sig,sig)
  
  # Get 30 random vectors
  Y<-rmvnorm(30,mu,Sigma)
  
  LDY<-matrix(nrow=ng,ncol=ng)
  for(i in 1:ng){
  for(j in 1:ng){ LDY[i,j]<-ldmvnorm( c(y[i],y[j]),mu,Sigma) }}
  
  plot(range(y),range(y),type="n", 
    xlab=expression(italic(y[1])),ylab=expression(italic(y[2])) )
  
  filledcontour(y,y,exp(LDY))
  points(Y,pch="x")
}

# dev.off()

```

```{r}
# OCT 20
#### Simulate inverse-Wishart matrix
# Gets a random vector from an inverset 
rinvwish<-function(n,nu0,iS0) 
{
  L0 <- chol(iS0) 
  S<-array( dim=c( dim(L0),n ) )
  for(i in 1:n) 
  {
    Z <- matrix(rnorm(nu0 * dim(L0)[1]), nu0, dim(iS0)[1]) %*% L0  
    S[,,i]<- solve(t(Z)%*%Z)
  }     
  S[,,1:n]
}
```


```{r}
setwd()
mu=t(mu0)
mu=t(mu)
mu
s0=matrix(c(1,0,0,2), nrow=2)
s0
rinvwish(1, mu, solve(s0))
```

```{r}

#### Simulate from the Wishart distribution
rwish<-function(n,nu0,S0)
{
  sS0 <- chol(S0)
  S<-array( dim=c( dim(S0),n ) )
  for(i in 1:n)
  {
    Z <- matrix(rnorm(nu0 * dim(S0)[1]), nu0, dim(S0)[1]) %*% sS0
    S[,,i]<- t(Z)%*%Z
  }
  S[,,1:n]
}
```


```{r}
#OCT 20
# IN HOMEWORK
# Look at Bayesian Inference for Multivariate normal model
#### Reading comprehension
setwd("/Users/masonnakamura/Downloads")
load("/Users/masonnakamura/Downloads/reading.RData")
Y<-reading
 
mu0<-c(50,50)
L0<-matrix( c(625,312.5,312.5,625),nrow=2,ncol=2)

nu0<-4
S0<-matrix( c(625,312.5,312.5,625),nrow=2,ncol=2)

n<-dim(Y)[1] ; ybar<-apply(Y,2,mean)
Sigma<-cov(Y) ; THETA<-SIGMA<-NULL
YS<-NULL
set.seed(1)

for(s in 1:5000) 
{
 
  ###update theta
  Ln<-solve( solve(L0) + n*solve(Sigma) ) # STEP 1
  mun<-Ln%*%( solve(L0)%*%mu0 + n*solve(Sigma)%*%ybar ) # STEP 1
  theta<-rmvnorm(1,mun,Ln)  # STEP 2
  ### 
   
  ###update Sigma
  Sn<- S0 + ( t(Y)-c(theta) )%*%t( t(Y)-c(theta) )  # STEP 3
  Sigma<-solve( rwish(1, nu0+n, solve(Sn)) ) # STEP 4
  ###

  ###
  YS<-rbind(YS,rmvnorm(1,theta,Sigma)) 
  ###

  ### save results 
  THETA<-rbind(THETA,theta) ; SIGMA<-rbind(SIGMA,c(Sigma))
  ###
  cat(s,round(theta,2),round(c(Sigma),2),"\n")
}
```

```{r}
# From above
quantile(  SIGMA[,2]/sqrt(SIGMA[,1]*SIGMA[,4]), prob=c(.025,.5,.975) )
quantile(   THETA[,2]-THETA[,1], prob=c(.025,.5,.975) )
mean( THETA[,2]-THETA[,1])
mean( THETA[,2]>THETA[,1]) 
mean(YS[,2]>YS[,1])
```

```{r}
#### Figure 7.2 
# pdf("fig7_2.pdf",family="Times",height=3.5,width=7)

# par(mfrow=c(1,2),mgp=c(1.75,.75,0),mar=c(3,3,1,1))
require("ash")

# This countour is a generalization of the confindence interval (here, its two dimensional)

plot.hdr2d(THETA,xlab=expression(theta[1]),ylab=expression(theta[2]) )
abline(0,1)
```

plot.hdr2d(YS,xlab=expression(italic(y[1])),ylab=expression(italic(y[2])), 
     xlim=c(0,100),ylim=c(0,100) )
points(Y[,1],Y[,2],pch=16,cex=.7)
abline(0,1)

dev.off()


#### Pima data 
data(Pima.tr)
Y0<-Pima.tr[,2:5]
Y<-Y0
n<-dim(Y)[1]
p<-dim(Y)[2]

set.seed(1)
O<-matrix(rbinom(n*p,1,.9),n,p)
Y[O==0]<-NA

#### Figure 7.3
pdf("fig7_3.pdf",family="Times", height=6,width=6)
par(mar=c(1,1,.5,.5)*1.75,mfrow=c(p,p),mgp=c(1.75,.75,0))
for(j1 in 1:p) {
for(j2 in 1:p) {
 if(j1==j2){hist(Y[,j1],main="");mtext(colnames(Y)[j1],side=3,line=-.1,cex=.7)}
  if(j1!=j2) { plot(Y[,j1],Y[,j2],xlab="",ylab="",pch=16,cex=.7)} 
                }}
dev.off()


####

## prior parameters
p<-dim(Y)[2]
mu0<-c(120,64,26,26)
sd0<-(mu0/2)
L0<-matrix(.1,p,p) ; diag(L0)<-1 ; L0<-L0*outer(sd0,sd0)
nu0<-p+2 ; S0<-L0
###

### starting values
Sigma<-S0
Y.full<-Y
for(j in 1:p)
{
  Y.full[is.na(Y.full[,j]),j]<-mean(Y.full[,j],na.rm=TRUE)
}
###


### Gibbs sampler
THETA<-SIGMA<-Y.MISS<-NULL
set.seed(1)

for(s in 1:1000)
{

  ###update theta
  ybar<-apply(Y.full,2,mean)
  Ln<-solve( solve(L0) + n*solve(Sigma) )
  mun<-Ln%*%( solve(L0)%*%mu0 + n*solve(Sigma)%*%ybar )
  theta<-rmvnorm(1,mun,Ln)
  ###
  
  ###update Sigma
  Sn<- S0 + ( t(Y.full)-c(theta) )%*%t( t(Y.full)-c(theta) )
#  Sigma<-rinvwish(1,nu0+n,solve(Sn))
  Sigma<-solve( rwish(1, nu0+n, solve(Sn)) )
  ###
  
  ###update missing data
  for(i in 1:n)
  { 
    b <- ( O[i,]==0 )
    a <- ( O[i,]==1 )
    iSa<- solve(Sigma[a,a])
    beta.j <- Sigma[b,a]%*%iSa
    s2.j   <- Sigma[b,b] - Sigma[b,a]%*%iSa%*%Sigma[a,b]
    theta.j<- theta[b] + beta.j%*%(t(Y.full[i,a])-theta[a])
    Y.full[i,b] <- rmvnorm(1,theta.j,s2.j )
  }
  
  ### save results
  THETA<-rbind(THETA,theta) ; SIGMA<-rbind(SIGMA,c(Sigma))
  Y.MISS<-rbind(Y.MISS, Y.full[O==0] )
  ###

  cat(s,theta,"\n")
}

#### Posterior mean and correlation matrix
apply(THETA,2,mean)

COR <- array( dim=c(p,p,1000) )
for(s in 1:1000)
{
  Sig<-matrix( SIGMA[s,] ,nrow=p,ncol=p)
  COR[,,s] <- Sig/sqrt( outer( diag(Sig),diag(Sig) ) )
}

apply(COR,c(1,2),mean)



##### Figure 7.5 
pdf("fig7_5.pdf",family="Times", height=7,width=7)
Y.true<-Y0
V<-matrix(1:p,nrow=n,ncol=p,byrow=TRUE)

v.miss<-V[O==0]
y.pred<-apply(Y.MISS,2,mean)
y.true<-Y.true[O==0]
par(mfrow=c(2,2),mar=c(3,3,1,1),mgp=c(1.75,.75,0))
for(j in 1:p){ plot(y.true[v.miss==j], y.pred[v.miss==j], 
          xlab=paste("true", colnames(Y.true)[j]), 
          ylab=paste("predictied", colnames(Y.true)[j]),pch=16 )
          abline(0,1)
  cat(j, mean( (y.true[v.miss==j]- y.pred[v.miss==j])^2),"\n") }
dev.off()



#### Convert SIGMA to an array of correlation parameters
COR<-array(dim=c(p,p,dim(SIGMA)[1]) )
for(s in 1:dim(SIGMA)[1]) {
Sig<-matrix( SIGMA[s,] ,nrow=p,ncol=p)
COR[,,s] <- Sig/sqrt(outer( diag(Sig),diag(Sig)))
                           }
colnames(COR)<-rownames(COR)<-colnames(Y)



#### Function for posterior quantile intervals for matrices
#### From the `sbgcop` package
plotci.sA<-function(sA, ylabs = colnames(sA[, , 1]), mgp = c(1.75, 0.75, 
    0)) 
{
    qA <- qM.sM(sA)
    p <- dim(qA)[1]
    tmp <- c(qA)
    tmp <- tmp[tmp != 1]
    par(mgp = mgp)
    for (j in 1:p) {
        plot(0, 0, type = "n", ylim = range(c(tmp), na.rm = TRUE), 
            xlim = c(1, p), ylab = ylabs[j], xaxt = "n", xlab = "")
        points((1:p)[-j], qA[j, -j, 2], pch = 16, cex = 0.6)
        segments(x0 = (1:p)[-j], y0 = qA[j, -j, 1], x1 = (1:p)[-j], 
            y1 = qA[j, -j, 3])
        abline(h = 0, col = "gray")
        abline(v = j, col = "gray")
    }
    axis(side = 1, at = 1:p, labels = colnames(qA[, , 1]), las = 2)
}

sR.sC<-function(sC) 
{
    p <- dim(sC)[1]
    s <- dim(sC)[3]
    sR <- array(dim = c(p, p, s))
    dimnames(sR) <- dimnames(sC)
    for (l in 1:s) {
        C <- sC[, , l]
        R <- C * NA
        for (j in 1:p) {
            R[j, -j] <- C[j, -j] %*% solve(C[-j, -j])
        }
        sR[, , l] <- R
    }
    sR
}

qM.sM<-function (sM, quantiles = c(0.025, 0.5, 0.975)) 
{
    p1 <- dim(sM)[1]
    p2 <- dim(sM)[2]
    s <- dim(sM)[3]
    qM <- array(dim = c(p1, p2, length(quantiles)))
    dimnames(qM) <- list(dimnames(sM)[[1]], dimnames(sM)[[2]], 
        paste(quantiles * 100, rep("% quantile", length(quantiles)), 
            sep = ""))
    for (l in 1:length(quantiles)) {
        qM[, , l] <- apply(sM, c(1, 2), quantile, prob = quantiles[l], 
            na.rm = TRUE)
    }
    qM
}




#### Figure 7.4 
pdf("fig7_4.pdf",height=6,width=6,family="Times")

par(mfcol=c(4,2),mar=c(1,2.75,1,1),mgp=c(1.75,.75,0),oma=c(1.5,0,0,0))
plotci.sA(COR)

REG<-sR.sC(COR)
plotci.sA(REG)
dev.off()

CQ<-apply(COR, c(1,2), quantile,prob=c(.025,.5,.975) )

round(CQ[1,,],2)
round(CQ[2,,],2)
round(CQ[3,,],2)

round(apply(COR,c(1,2),mean),2)



