---
title: "Bayesian hierarchical modeling"
output:
  pdf_document: default
  html_document:
    number_sections: yes
fontsize: 11pt
date: "2022-10-27"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
require(runjags)
require(coda)
crcblue <- "#2905a1"
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Installing the necessary packages

```{r, eval = FALSE}
# install.packages("devtools")
require(devtools)
# devtools::install_github("bayesball/ProbBayes")

require(ggplot2)
require(gridExtra)
require(ProbBayes)
require(tidyverse)
crcblue <- "#2905a1"
```



# Example: Korean Drama Ratings

## Ratings by Schedule

```{r message = FALSE}
dramadata = read.csv("https://foxweb.marist.edu/users/duy.nguyen2/KDramaData.csv", header=T)

KBSdrama = dramadata[dramadata$Producer==2,]
KBSdrama$Schedule = as.factor(KBSdrama$Schedule)
```

```{r echo = FALSE}
ggplot(KBSdrama, aes(Rating, color = Schedule)) +
  geom_density() + labs(title = "Density plot of ratings") + 
  xlim(0, 0.3) + theme_grey(base_size = 10, base_family = "") 
```


```{r}
table(KBSdrama$Schedule)
tapply(KBSdrama$Rating, KBSdrama$Schedule, summary)
tapply(KBSdrama$Rating, KBSdrama$Schedule, sd)
```

# MCMC simulation by JAGS

## JAGS Script for the Hierarchical Model

```{r message = FALSE}
modelString <-"
model {
## likelihood
for (i in 1:N){
  y[i] ~ dnorm(mu_j[schedule[i]], invsigma2)
}

## priors
for (j in 1:J){
mu_j[j] ~ dnorm(mu, invtau2)
}
invsigma2 ~ dgamma(a_g, b_g)
sigma <- sqrt(pow(invsigma2, -1))

## hyperpriors
mu ~ dnorm(mu0, 1/g0^2)
invtau2 ~ dgamma(a_t, b_t)
tau <- sqrt(pow(invtau2, -1))
}
"
```


- Pass the data and hyperparameter values to JAGS:

```{r message = FALSE}
y = KBSdrama$Rating   
print(y)
schedule = KBSdrama$Schedule  
print(schedule)
N = length(y)  
J = length(unique(schedule)) 

# for the replication of ~set.seed(123)
initsfunction <- function(chain){
  .RNG.seed <- c(1,2)[chain]
  .RNG.name <- c("base::Super-Duper",
                 "base::Wichmann-Hill")[chain]
  return(list(.RNG.seed=.RNG.seed,
              .RNG.name=.RNG.name))
}

the_data <- list("y" = y, "schedule" = schedule, "N" = N, "J" = J, 
                 "mu0" = 0.1, "g0" = 0.5, 
                 "a_t" = 1, "b_t" = 1,
                 "a_g" = 1, "b_g" = 1)
```

- Run the JAGS code for this model:

```{r message = FALSE, warning = FALSE}
posterior <- run.jags(modelString,
                      n.chains = 1,
                      data = the_data,
                      monitor = c("mu", "tau", "mu_j", "sigma"),
                      adapt = 1000,
                      burnin = 5000,
                      sample = 5000,
                      thin = 1, 
                      inits = initsfunction)
```

## JAGS Output of the Hierarchical Model

- Obtain posterior summaries of all parameters:

\vspace{3mm}

```{r message = FALSE, warning = FALSE}
summary(posterior)
# Since the means vary largely depending on the scheduling for the subgroups, the scheduling matters
# Note that the movie rating is below 0 for the lower95 
# Approach 1 --> we can truncate/throw away anything on the domain of the pdf that is < 0
# Approach 2 --> we can use the log normal distribution to avoid a domain < 0 on the pdf
```

```{r message = FALSE}
plot(posterior, vars = "mu_j[1]")
# => mu_j[1] converges in top left plot
# NOTE: if the graph in the top left plot does not seem to converge, then run more iterations until convergence

# In plot bottom right, if the correlation is unstable and not 0 as seen below, then we need more iterations
```
```{r message = FALSE}
plot(posterior, vars = "mu_j[2]")
```

```{r message = FALSE}
plot(posterior, vars = "tau")
```

## Shrinkage/Pooling Effects

```{r message = FALSE, warning = FALSE}
Ind_Stats = as.data.frame(matrix(NA, J, 2))
names(Ind_Stats) = c("mean", "sd")
for (j in 1:J){
  Ind_Stats[j, ] = c(mean(KBSdrama$Rating[KBSdrama$Schedule == j]), 
                     sd(KBSdrama$Rating[KBSdrama$Schedule == j]))
}

Post_Means <- summary(posterior)[, 4]

Means1 <- data.frame(Type = "Sample", Mean = Ind_Stats$mean)
Means2 <- data.frame(Type = "Hierarchical", Mean =
                       Post_Means[3:(4 + J - 2)])

Means1$Title <- c("Schedule 1", "Schedule 2", "Schedule 3",
                  "Schedule 4")
Means2$Title <- c("Schedule 1", "Schedule 2", "Schedule 3",
                  "Schedule 4")
```


```{r message = FALSE}
ggplot(rbind(Means1, Means2), aes(Type, Mean, group=Title)) +
  geom_line(color = crcblue) + geom_point() +
  annotate(geom = "text", x = 0.75,
           y = Means1$Mean + c(0.01, 0.01, 0.01, -0.01), 
           size = 3, label = Means1$Title) + increasefont(Size = 10)
```


## Sources of Variability
## We need the \texttt{coda} R package

```{r eval = FALSE}
install.packages("coda")
```

```{r message = FALSE, warning = FALSE}
require(coda)
tau_draws <- as.mcmc(posterior, vars = "tau")
sigma_draws <- as.mcmc(posterior, vars = "sigma")
R <- tau_draws^2/(tau_draws^2 + sigma_draws^2)

df <- as.data.frame(R)

quantile(R, c(0.025, 0.975))
```


```{r message = FALSE}
ggplot(df, aes(x=R)) + geom_density() +
  labs(title="Density of R") +
  theme(plot.title = element_text(size=15)) +
  theme(axis.title = element_text(size=15))
```


# In cl





# Recap